\section{Data Description}

The dataset is obtained from the UC Irvine Machine Learning Repository \cite{b2}. The dataset titled "MSNBC.com Anonymous Web Data Data Set" \cite{b2}, describes the page visits of users who visited msnbc.com on September 28, 1999. Visits are recorded at the level of URL category and are recorded in time order. The data comes from Internet Information Server (IIS) logs for msnbc.com and news-related portions of msn.com for the entire day of September, 28, 1999 (Pacific Standard Time). Each sequence in the dataset corresponds to page views of a user during that twenty-four hour period. Each event in the sequence corresponds to a user's request for a page. Requests are not recorded at the finest level of detail---that is, at the level of URL, but rather, they are recorded at the level of page category (as determined by a site administrator). The categories are ``frontpage", ``news", ``tech", ``local", ``opinion",``on-air", ``misc", ``weather", ``health", ``living", ``business", ``sports", ``summary", ``bbs" (bulletin board service), ``travel", ``msn-news", and ``msn-sports".

The dataset contains 989818 rows of sequences with each event in the sequence being labelled using integers. Each integer corresponds to a category, for example, ``frontpage" is represented as 1, ``news" with 2, ``tech" with 3, etc. The average sequence length is 4.7 but ranges from a length of 1 to a maximum length of 14795. Each sequence is a list of numbers and hence all events in all sequences have only a single element.

This dataset is perfectly structured to perform sequential pattern mining and the resulting frequent patterns obtained from SPADE will give us the sequences in which users view different page categories on msnbc.com